%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
The introduction should be approximately 1–3 pages in length, and should contain the following information:

Problem statement: Make a concise statement of the problem, ideally in a few sentences, but no more than a paragraph. For example, try to complete this statement: “The sponsor desires that ... (insert goals of the project) ... subject to the following criteria: ... (insert numbered list).” These goals and criteria help to define the scope of work and the deliverables.
Background or Related Work: State who else has worked on this problem or similar problems (you should do most of your citations here). For applied projects, provide information on other existing programs which will use your program.
Objectives: The objectives are a battle plan for the project. They are a breakdown of steps or accomplishments that must be completed to achieve the project goals.

\section{Problem Statement}
Current methods of vocal communication relies on audible speech or text-based inputs. This creates barriers for individuals with speech disabilities and vocal impairments. 
Surface electromyography (sEMG) offers a way to capture muscle activity during vocalizations
 These signals can be translated into meaningful speech representations, such as the International Phonetic Alphabet (IPA), yet it remains a challenge.By developing a Silent Speech Interface [1] that accurately maps sEMG signals from facial and neck muscles groups involved in vocalization to IPA symbols we would be able to provide real-time vocal communication for non-speaking individuals or enhance existing silent speech technologies. The primary challenge lies in decoding complex muscle signal patterns, ensuring a consistent and high accuracy across a diverse set of users, and creating a system robust enough for real-world applications. This gap in accessibility and communication necessitates a need for an innovative solution combining biomechanics, linguistics, and machine learning.


\section{Background or Related Work}
State who else has worked on this problem or similar problems (you should do most of your citations here). For applied projects, provide information on other existing programs which will use your program.

Describe what systems already exist and why they are inadequate. 

\section{Objectives}
The objectives are a battle plan for the project. They are a breakdown of steps or accomplishments that must be completed to achieve the project goals.

\section{Our approach}
Describe the team's approach for developing a system at a high level. Why will your work result in a system that is different / better than existing solutions? 
